---
title: "Linear Mixed Effects Models - 2 - results"
author: "Kiri Koppelgaard"
date: "September 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language development in Autism Spectrum Disorder (ASD)

Background: Autism Spectrum Disorder is often related to language impairment. However, this phenomenon has not been empirically traced in detail:
i) relying on actual naturalistic language production, 
ii) over extended periods of time.

We therefore videotaped circa 30 kids with ASD and circa 30 comparison kids (matched by linguistic performance at visit 1) for ca. 30 minutes of naturalistic interactions with a parent. We repeated the data collection 6 times per kid, with 4 months between each visit. We transcribed the data and counted: 
i) the amount of words that each kid uses in each video. Same for the parent.
ii) the amount of unique words that each kid uses in each video. Same for the parent.
iii) the amount of morphemes per utterance (Mean Length of Utterance) displayed by each child in each video. Same for the parent. 

This data is in the file you prepared in the previous class. 

NB. A few children have been excluded from your datasets. We will be using them next week to evaluate how good your models are in assessing the linguistic development in new participants.

We then want to test the language trajectory of child and parent over time.

This RMarkdown file is structured in the following way:

1. The exercises: read them carefully. Under each exercise you will have to write your answers, once you have written and run the code. This is the part that you have to directly send to the teachers.
2. An (optional) guided template full of hints for writing the code to solve the exercises. Fill in the code and the paragraphs as required. Then report your results under the exercise part.
3. In exercise 4 you will be asked to create the best possible model of language development in TD and ASD children, picking and choosing whatever additional variables you want from the dataset. Next time, the models produced by the different groups will compete against each other to see who can produce the best model, so choose carefully!

You will have to have a github repository for the code and send the answers to Malte and Riccardo without code (but a link to your github/gitlab repository). This way we can check your code, but you are also forced to figure out how to report your analyses :-)

Remember to submit only your findings, and not just the code. To do this you can either
- Write your answers in a separate document
- Write your answers in the template, but tell rstudio not to print the code chunks when you knit it with the chunk option include=FALSE

```{r, include=FALSE}
# with include=FALSE, the code will still run
x = 1
# but it will not print anything
print(x)
```

```{r, echo = FALSE}
# if you want to hide the code but show the output (a plot for instance)
# you can use echo = FALSE
print(x+10)
```

To re-iterate one more time: Hand in a document with your findings but without code (html or pdf or word) and a link to your github/gitlab with your Rmd file.

N.B. The following lines are a summary of the questions to be answered, the step-by-step instructions and tips are in the template.

```{r, include = FALSE}
library(pacman)
p_load(tidyverse, stringr, lme4, nlme, growthcurver, dplyr, reshape2, purr)
setwd("~/Cognitive Science/3. Semester/Experimental Methods 3/Portfolios/Data/")
autism_data <- read.csv("autism_data.csv")
```

## Exercise 1) Preliminary Data Exploration

Describe the participant samples in the dataset (e.g. by diagnosis, age, etc.). Do you think the two groups are well balanced? If not, what do you think was the reason?

```{r, echo = FALSE}
#Diagnosis
ggplot(autism_data, aes(Diagnosis)) +
  geom_bar(aes(fill=Diagnosis))
```

Equal distribution of dignosis (cirka) - but not representive for the rest of the population. 

```{r, echo = FALSE}
ggplot(autism_data, aes(Diagnosis, Age)) + 
  geom_boxplot()
```

With diagnosis ASD often older - though spread also larger. If the model used to predict mlu used age rather than visit - one might imagine the effect to be larger, since the children with ASD are further along in their development. 

```{r,  echo = FALSE}
#Demographics
ggplot(autism_data, aes(Gender)) +
  geom_bar(aes(fill=Diagnosis), position = "dodge")
  

```
The distribution of gender is unbalanced with a high amount of females. This seems counterintutive, since previous research have shown that ASD is mostly 

```{r, echo = FALSE}

autism_data$Ethnicity[34] <- "Bangladeshi" 

ggplot(autism_data, aes(Ethnicity)) +
  geom_bar(aes(fill=Ethnicity)) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(autism_data, aes(Diagnosis)) +
  geom_bar(aes(fill=Gender, position = "dodge")) + 
             facet_wrap(~Ethnicity) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
A lot more white - if you want to say something about ethnicity - then a problem

If compare to diagnosis - not all the whites are normal. 

```{r, echo = FALSE}

# ADOS 
ggplot(autism_data, aes(Diagnosis, ADOS1)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 

```

Instrument for assesing ASD - so makes sense higher for children with ASD - the point is the measure works 

```{r, echo = FALSE}
# VerbalIQ
ggplot(autism_data, aes(Diagnosis, verbalIQ1)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 

# nonverbalIQ1
ggplot(autism_data, aes(Diagnosis, nonverbalIQ1)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 


```
 Verbal IQ seems to relate to the diagnosis - if you are typical development you have a higher bverbal iq - this is interesting in relation to non-verbal IQ - see next plot 

Higher for children with ASD interesting in relation to typically developed. 

```{r}

#tokens
ggplot(autism_data, aes(Diagnosis, tokens_CHI)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 
#types
ggplot(autism_data, aes(Diagnosis, types_CHI)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 

```

Children with ASD usually have smaller vocabulary and generally speak less. 



```{r, echo = FALSE}
#child mlu
ggplot(autism_data, aes(Diagnosis, CHI_MLU)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 
#parent mlu
ggplot(autism_data, aes(Diagnosis, MOT_MLU)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 
```
Differs from tokens by telling how much children talk in every utterance. Thus, ASD tends to talk less, use less different words and make shorter utterances. 

Generally mother to children with ASD aso talk a bit less - interesting whether this is an effect of the children speaking less or if the children speak less because their mothers do. 

### Exercise 2) Children learning language: the effects of time and ASD
Describe linguistic development in TD and ASD children in terms of Mean Length of Utterance (MLU)?

We hypothesise that child MLU is predicted by visit and diagnosis. 

First we tried to describe the correlation with linear regression, however comparing the linear model with a null model, where mean length utterance only is predicted by time. This appears strange looking at the plots we described in the beginng, where diagnosis seem to have an effect of MLU. Therefore we try to fit a growth curve. 


We used R (R Core Team, 2012) and lme4 (Bates, Maechler & Bolker, 2012) to perform a CUBIC/QUADRATIC? mixed effects analysis of the relationship between the children's mean length of utterance as it evolves over time and their diagnosis. As fixed effects, we entered diagnosis (ASD; autism spectrum disorder and TP; typically developing) and visit (without interaction term) into the model. As random effects, we had an intercept for subjects, as well as a by‐subject random slope for the effect of visit. P‐value was obtained by likelihood ratio tests of the full model with the effect in question against the model without the effect in question; children's diagnosis and number of visit affected mean length of utterance (χ2(4)=63.82,	 p<0.05),	mean length of utterance 	was extended by	about	0.41 words per visit	(SE= 0.049) and was extended by	about 0.11 (SE=0.12) for typically developing children. 

Thus, childrens mean length utterane was significantly predicted by visit and diagnosis. The development of mean length of utterance for the children is shown in the plot below. 

```{r, echo = FALSE}
ggplot(autism_data, aes(x=VISIT, y = CHI_MLU)) + geom_point(alpha=0.5, aes(color = Diagnosis)) + geom_smooth(method=lm,formula=y~poly(x,3), aes(color=Diagnosis))+ theme_bw()
```


### Exercise 3) Child directed speech as a moving target
Describe how parental use of language changes over time in terms of MLU. What do you think is going on?



### Exercise 4) Looking into "individual differences" (demographic, clinical or cognitive profiles)
The dataset contains some additional variables characterizing the kids’ cognitive and clinical profile: ADOS (autism severity), MSEL EL (Expressive Language, that is, verbal IQ, or linguistic skills at first visit as assessed by a psychologist using Mullen Scales of Early Learning), MSEL VR (Visual Reception, used as a proxy for non verbal IQ at first visit), Age, Gender, Ethnicity. Would it make sense to add any of them to your model of linguistic trajectories? Create the best possible model (the one that best explain the data, with MLU as outcome). Next time your model will be tested on new participants, and we will proclaim a winner. Describe your strategy to select the best models (how did you choose the variables to include?) and send the code to Riccardo and Celine.

### [OPTIONAL] Exercise 5) Comment on how the three linguistic variables measure linguistic performance (the so-called "construct validity" of the measures). Do they express the same variance?



