---
title: "Assignment 1 - Language Development in ASD - part 2"
author: "Kiri Koppelgaard"
date: "September 12, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# you could optionally set include = TRUE globally while working on the exercise and then just change it to include = FALSE before you hand in
# knitr::opts_chunk$set(include = FALSE)
options(warn=-1) #turning of warnings

```
The following is based on the preprocessing from the first part of assignment 1: Language Development in ASD, which can be found on GitHub: https://github.com/KiriKoppelgaard/Portfolios/blob/master/Linear%20Mixed%20Effects%20models%20-%201.Rmd

We will try to answer three questions:
    
- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries, include = FALSE}
library(pacman)
p_load(tidyverse, stringr, lme4, nlme, growthcurver, dplyr, reshape2, purr, effects, MuMIn, ggbeeswarm)

```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

If your're in a project, just put the data in the project folder and you're good to go! (Or make a data subfolder to keep it tidy around here)
```{r Load Data, include = FALSE}
setwd("~/Cognitive Science/3. Semester/Experimental Methods 3/Portfolios/Data/")
autism_data <- read.csv("autism_data.csv")
```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

REPONSE: 
Generally, the sample of participants are proportionally distributed according to the most relevant variables. This enables application of statistical tools which will produce relatively reliable  predictions of the mean length of utterance of children predicted by time and their diagnosis. However, there are several issues I would like to point to.  

As seen below the distribution of typically developed children and children with ASD is relatively equal
```{r, echo = FALSE}
#Making grouped data
grouped_data <- autism_data %>%
  filter(Gender %in% c("F", "M")) %>%
  group_by(ANO_SUBJ, Gender) %>%
  summarise(counts = n()) 

autism_data$Ethnicity[34] <- "Bangladeshi" 

grouped_data2 <- autism_data %>%
  group_by(ANO_SUBJ, Ethnicity) %>%
  summarise(counts = n()) 
grouped_data2 <- na.omit(full_join(grouped_data, grouped_data2))


grouped_data3 <- autism_data %>%
  group_by(ANO_SUBJ, Diagnosis) %>%
  summarise(counts = n())

grouped_data3 <- na.omit(full_join(grouped_data3, grouped_data2))

grouped_data4 <- autism_data %>%
  group_by(ANO_SUBJ) %>%
  summarise(Age = mean(Age))

grouped_data4 <- na.omit(full_join(grouped_data3, grouped_data4))

#Diagnosis
ggplot(grouped_data3, aes(Diagnosis)) +
  geom_bar(stat = "count", aes(fill=Diagnosis))


```


A comment to be made in relation to the distribution of typically developing children compared to children with ASD is that this proportion does not map onto the population. However, since the variable we wish to measure is the mean length of utterance in relation to time and diagnosis, the distribution of diagnosis play little role.

A comment can be made in relation to the distribution of age of in relation to diagnosis. As seen in the plot below, the sample has a disproportionate age distribution. It appears that the sampled children with ASD are slightly older than typically developed children. Since we use visits as a predictor of mean length of utterance rather than age, this might have the effect that the sampled children with ASD will be further along in their development without us taking into account that they were older in the first place. 

```{r, echo = FALSE}
#Plotting age
ggplot(grouped_data4, aes(Diagnosis, Age)) + 
  geom_violin(aes(fill = Diagnosis)) +
  stat_summary(fun.y=median, geom="point", size=2, color="red") +
  labs(y = "Age in months") + geom_boxplot(width=0.1)+ 
  stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red")
```


The distribution of gender is pretty skewed, however, this corresponds with the population, since mostly males are diagnosed with ASD. 

```{r, echo = FALSE}
#Plotting gender
ggplot(grouped_data, aes(x = Gender)) +
  geom_bar(stat = "count", aes(fill=Gender))
```

The distribution of ethnicity shows a high amount of whites compared to other ethnicities. However, since we have little interest in the effect of ethnicity on mean length of utterance, this is not too alarming. Especially, since the distribution of diagnosis within the catgories of ethnicity show a relatively equal distribution. 

```{r descriptive stats, echo = FALSE}
#Plotting ethnicity
ggplot(grouped_data2, aes(Ethnicity)) +
  geom_bar(stat = "count", aes(fill=Ethnicity)) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(grouped_data3, aes(Diagnosis)) +
   geom_bar(stat = "count", aes(fill=Gender), position = position_dodge(0.8)) + 
              facet_wrap(~Ethnicity) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


The following plots are highly relevant for assessing, whether the measures of diagnosing ASD are any good. 
The plot below visualize that the mean ADOS value (an instrument for diagnosing and assessing autism) appears higher for children with ASD than typically developing children - thus, the measure works. However, the diagnosis of the children is probably based on this measure, therefore it might in reality not tell us much. 

```{r, echo = FALSE}

# ADOS 
ggplot(autism_data, aes(Diagnosis, ADOS1)) + 
  geom_quasirandom(alpha = 0.5, colour = "palegreen4") + labs(x = "Diagnosis", y = "Parameter value", title = "ADOS") + geom_boxplot(alpha = 0.5)

```

The linguistic intelligence of the children are measured using a verbal IQ measure. The verbal IQ would we  expect to be lower for children with ASD, since we hypothesise children with ASD in general use fewer words, also unique words and speak less in coherence. The plot by apperance supports this hypothesis, since typically developing childern has a higher verbal IQ. 

```{r, echo = FALSE}
# VerbalIQ
ggplot(autism_data, aes(Diagnosis, verbalIQ1)) + 
  geom_quasirandom(alpha = 0.5, colour = "palegreen4") + labs(x = "Diagnosis", y = "Parameter value", title = "VerbalIQ") + geom_boxplot(alpha = 0.5)

```

The following plots show a distribution of how many words the sample's children utter and how many unique words they utter. In line with our beliefs, children with ASD appears to have smaller vocabulary and generally speak less.

```{r, echo = FALSE}

#tokens
ggplot(autism_data, aes(Diagnosis, tokens_CHI)) + 
  geom_quasirandom(alpha = 0.5, colour = "palegreen4") + labs(x = "Diagnosis", y = "Parameter value", title = "Tokens") + geom_boxplot(alpha = 0.5)

#types
ggplot(autism_data, aes(Diagnosis, types_CHI)) + 
  geom_quasirandom(alpha = 0.5, colour = "palegreen4") + labs(x = "Diagnosis", y = "Parameter value", title = "Types") + geom_boxplot(alpha = 0.5)

```

Lastly reported are bar plots of the children and their mother's s mean length of utterance. By appearance this shows that children with ASD tend to talk less in coherence - a pattern which is reflected in their mothers. Generally mothers to children with ASD also talk less in coherence. An interesting matter to investigate is whether this is an effect of the children speaking less or if the children speak less because their mothers do. 


```{r, echo = FALSE}
#child mlu
ggplot(autism_data, aes(Diagnosis, CHI_MLU)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 
#parent mlu
ggplot(autism_data, aes(Diagnosis, MOT_MLU)) + 
         stat_summary(fun.y = mean, geom = "bar", fill = "White", colour = "Black") + 
                        stat_summary(fun.data = mean_cl_normal, geom = "pointrange") 
```




## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r ex2, include = FALSE}

model = lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|ANO_SUBJ), data = autism_data, REML = FALSE)

summary(model)
```

How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}
#Buidling a baseline model
intercept_model =lmer(CHI_MLU ~ 1+ (1+VISIT|ANO_SUBJ), data = autism_data,
REML=FALSE)# creating the intercept model

#Buidling a null model
null_model =lmer(CHI_MLU ~ VISIT + (1+VISIT|ANO_SUBJ), data = autism_data,
REML=FALSE)# creating the null model

#Comparing the models using anova()
anova(null_model,intercept_model, model)

```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r ex2 growth curve, include = FALSE}
#building quadratic model
model2 = lmer(CHI_MLU ~ VISIT + Diagnosis + I(VISIT^2)+(1+VISIT+ I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)

#building interaction model
model3 = lmer(CHI_MLU ~ VISIT * Diagnosis+(1+VISIT+ I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)

#building interaction quadratic model
model4 = lmer(CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2)*Diagnosis+(1+VISIT+ I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)

summary(model4)

# #Building cubic model
# autism_data$VISIT1 <- autism_data$VISIT/6 #trying to scale VISIT to get it to converge
# model5 = lmer(CHI_MLU ~ Diagnosis *VISIT + Diagnosis *I(VISIT^2) +Diagnosis *I(VISIT^3)+(1+VISIT+ I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)
# #FAILS TO CONVERGE, removed third order random effects, but I do not trust it.


#testing them against each other
anova(model2, model3, model4)


#testing it against the quadratic model
anova(intercept_model, null_model, model4)

summary(model4)

r.squaredGLMM(model4)

```

Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your best model's predictions

RESPONSE: 
Linguistic development of childrens' MLU is affected by the interaction between visit (time) and diagnosis (MLU ~ visit * diagnosis + visit^2 * diagnosis).  

Using lme4 to perform a quadratic mixed effects analysis of the relationship between the children's mean length of utterance as it evolves over time and their diagnosis. 

As fixed effects, we entered the interaction effect between diagnosis (ASD; autism spectrum disorder and TD; typically developing) and visit and the interaction effect between the quadratic effect of visit and diagnosis (omitting their main effects) into the model. 

As random effects, we had an intercept for subjects, as well as a by-subject random slope for the effect of visit - both linear and quadratic. 

P-value was obtained by likelihood ratio tests of the full model with the effect in question against the model without the effect in question; children's diagnosis and number of visit affected mean length of utterance (X^2(7)=100.973,	 p<0.05), we observe a significant increase for the interaction between diagnosis and the linear effect of visit (0.48, SE = 0.14, p < 0.05) accompanied by a nonsignificant quadratic component of the growth curve for the interaction between visit^2 and diagnosis(-0.03, SE = 0.019, p > 0.05), which if significant indicates a slow down of the increase over time. As can be observed in figure it appears  that there is a slow down of the increase over time, this is however not significant. The marginal r-squared i.e. how much variance is explained by the systematic components of your model is 0.37.   

As seen in the second plot ASD children's development is less homogenous than TD children's, which makes it harder to predict. 


```{r, echo = FALSE}
#plotting the results

ggplot(autism_data, aes(x=VISIT, y = CHI_MLU)) + geom_point(alpha=0.5, aes(color = Diagnosis)) + geom_smooth(method=lm,formula=y~poly(x,2), aes(color=Diagnosis))+ theme_bw()

autism_data$ANO_SUBJ <- as.character(autism_data$ANO_SUBJ)

ggplot(autism_data,aes(x=VISIT,y=CHI_MLU,colour=ANO_SUBJ)) + geom_point(show.legend = FALSE) + geom_smooth(method="lm",formula=y~poly(x,2),aes(colour=ANO_SUBJ),se=FALSE, show.legend = FALSE) + facet_wrap(~Diagnosis)

#If I want to plot this without the interaction term, what do I do? 

#Plotting random effects
# x1 <- as.data.frame(effect("VISIT:Diagnosis", model4))
# x2 <- as.data.frame(effect("Diagnosis:I(VISIT^2)", model4))
# 
# ggplot(x1, aes(VISIT, fit, color=Diagnosis)) + geom_point() + geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4) + theme_bw(base_size=12)
# 
# ggplot(x2, aes(I(VISIT^2), fit, color=Diagnosis)) + geom_point() + geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4) + theme_bw(base_size=12)

```



## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}

#building a linear model
model6 = lmer(MOT_MLU ~ VISIT + Diagnosis + (1+VISIT|ANO_SUBJ), data = autism_data, REML = FALSE)

summary(model6)

#building quadratic model
model7 = lmer(MOT_MLU ~ VISIT + Diagnosis + I(VISIT^2)+(1+VISIT+ I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)


#Building cubic model
model8 = lmer(MOT_MLU ~ Diagnosis + VISIT  + I(VISIT^2) +I(VISIT^3)+(1+VISIT|ANO_SUBJ), data = autism_data, REML = FALSE)

#Building interaction quadratic model
model9 = lmer(MOT_MLU ~ Diagnosis*VISIT  + Diagnosis*I(VISIT^2)+(1+VISIT+I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)

#Building cubic model
model10 = lmer(MOT_MLU ~ Diagnosis * VISIT  + Diagnosis* I(VISIT^2) +Diagnosis*I(VISIT^3)+(1+VISIT+I(VISIT^2)+I(VISIT^3)|ANO_SUBJ), data = autism_data, REML = FALSE)

#Buidling a baseline model
intercept_model1 =lmer(MOT_MLU ~ 1+ (1+VISIT|ANO_SUBJ), data = autism_data,
REML=FALSE)# creating the intercept model

#Buidling a null model
null_model1 =lmer(MOT_MLU ~ VISIT + (1+VISIT|ANO_SUBJ), data = autism_data,
REML=FALSE)# creating the null model

#Comparing the models using anova()
anova(model6,model7, model8, model9, model10)

#Comparing the models using anova()
anova(null_model1,intercept_model1, model7)

summary(model7)

r.squaredGLMM(model7)
```

RESPONSE: The parental use of language is affected by visit and diagnosis (MLU ~ visit + visit^2 + diagnosis) in line with their children.  

Using lme4 to perform a quadratic mixed effects analysis of the relationship between the mothers' mean length of utterance as it evolves over time and their childrens diagnosis. 

As fixed effects, we entered the main effect of diagnosis (ASD; autism spectrum disorder and TP; typically developing) visit and the quadratic effect of visit into the model. 

As random effects, we had an intercept for subjects, as well as a by-subject random slope for the effect of visit - both linear and quadratic. 

P-value was obtained by likelihood ratio tests of the full model with the effect in question against the model without the effect in question; children's diagnosis and number of visits significantly affected parental mean length of utterance (x^22(5)=31.928,	 p<0.05),	 we observe a significant linear increase for the effect of diagnosis (0.49, SE = 0.11, p > 0.05) and for the effect of visit (0.28, SE = 0.06, p < 0.05) accompanied by a significant quadratic component of the growth curve for the effect of visit (-0.02, SE = 0.008, p < 0.05). As can be observed in the figure below, this indicates a slow down of the increase over time. The marginal r-squared is 0.23. 


```{r, echo= FALSE}
ggplot(autism_data, aes(x=VISIT, y = MOT_MLU)) + geom_point(alpha=0.5, aes(color = Diagnosis)) + geom_smooth(method=lm,formula=y~poly(x,3), aes(color=Diagnosis))+ theme_bw()
```

The similar effect of shorter mean length of utterances for parents there have children with ASD is probably a result of an adjustment to the child's progress. 

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Malte (au540041@post.au.dk).


```{r ex4, include = FALSE}

#Adding variables to find best model to desribe development of CHI_MLU
model11 = lmer(CHI_MLU ~ VISIT + Diagnosis + verbalIQ1 + Gender +(1+VISIT|ANO_SUBJ)+ (1|Ethnicity), data = autism_data, REML = FALSE)

model12 = lmer(CHI_MLU ~ VISIT + Diagnosis + verbalIQ1 + nonverbalIQ1 +  Gender +(1+VISIT|ANO_SUBJ)+  (1|Ethnicity), data = autism_data, REML = FALSE)

model13 = lmer(CHI_MLU ~ VISIT + Diagnosis + ADOS1 + Gender + (1+VISIT|ANO_SUBJ)+  (1|Ethnicity), data = autism_data, REML = FALSE)

model14 = lmer(CHI_MLU ~ VISIT + Diagnosis + ADOS1 + verbalIQ1 + nonverbalIQ1 + (1+VISIT|ANO_SUBJ)+  (1|Ethnicity), data = autism_data, REML = FALSE)

model15 = lmer(CHI_MLU ~ VISIT + Diagnosis  + ADOS1 + verbalIQ1 + nonverbalIQ1 + Gender+  I(VISIT^2)+(1+VISIT+ I(VISIT^2)|ANO_SUBJ)+ (1|Ethnicity), data = autism_data, REML = FALSE)

model16 = lmer(CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2)*Diagnosis+ verbalIQ1 + nonverbalIQ1 +(1+VISIT+ I(VISIT^2)|ANO_SUBJ), data = autism_data, REML = FALSE)



#testing new models against the quadratic model - the best model so far
anova(model4, model11, model12, model13, model14, model16, model15)

#...and the winner is!
summary(model16)

#Comparing the models using anova()
anova(null_model1,intercept_model1, model16)

r.squaredGLMM(model16)
```

REPONSE: 
The best model found by a serious amount of p-hacking was: Childrens mean length of utterance ~ VISIT * Diagnosis + VISIT^2*Diagnosis+ verbal IQ + nonverbal IQ +(1+VISIT+ VISIT^2|ANO_SUBJ)

As fixed effects, we entered the interaction between diagnosis and the linear effect of visit, the interaction between the quadratic component of visit and diagnosis (disregarding their main effects), verbal IQ and nonverbal IQ into the model. As random effects, we had a by subject random slope for the effect of visit - both linear and quadratic. 

P-value was obtained by likelihood ratio tests of the full model with the effect in question against the model without the effect in question; children's diagnosis, verbal IQ, nonverbal IQ and number of visits affected the mean length of utterance (X^2(9)=65.677,	 p<0.05),	 we observe a significant increase for the effect of the interaction between diagnosis and the linear component of visit (0.47, SE = 0.13, p < 0.05) accompanied by a non-significant effect of the interaction between the quadratic component of the growth curve for the effect of visit and diagnosis(-0.03, SE = 0.019, p > 0.05).A significant linear increase is observed for the effect of verbal IQ (0.06, SE = 0.009, p < 0.05) along with a non-significant effect of nonverbal IQ (-0.005, SE = 0.01, p > 0.05). The marginal r-squared is 0.6. 

For code see GitHub: 

